Estimating duration using survival analysis

### 2024-03-22 cis-perfect-testing.tex
Updates from Overleaf

% \section{Definitions}
% \todo[inline]{These will be defined in the introductory chapters, but are here so that this draft is understandable}
% \begin{itemize}
% \item
%   We assume that, for all episodes $i$, $D_i$ is iid and independent
%   of the time of the infection. Define the survival function
%   $\prob(D_i \geq t \mid B_{j(i)} = b, \vec{\theta}) = \prob(D_i \geq t \mid \vec{\theta}) = S_{\vec{\theta}}(t)$,
%   where $\vec{\theta}$ are the parameters controlling the survival
%   distribution (the discussion in this document is valid regardless of
%   the model specified for $S_{\vec{\theta}}$ and hence we consider $\vec{\theta}$
%   as an arbitrary vector of parameters).
% \item
%   The beginning of episode $i$ is known to occur in the interval
%   $[l_{j(i)}^{(b)}, r_{j(i)}^{(b)}]$, and similarly for the end of the infection
%   in $[l_{j(i)}^{(e)}, r_{j(i)}^{(e)}]$.
% \end{itemize}

@@

Doubly interval censored and truncated data are common when assessing the survival time of bird and insect nests~\autocite{heiseyABCs}.
\textcite{heiseyModelling}, building on long-standing methodology~\autocite{dempsterMaximum,turnbullEmpirical}, develops a likelihood-based inference framework, allowing for arbitrary truncation patterns and double interval censoring.
They apply it to a situation where the visit times (possible detection times) are common to all the individuals, simplifying the likelihood.
I adapt this framework for CIS, when each individual has their own visit schedule (see section \cref{perf-test:sec:model}).

@@

The model will be used to analyse on the 4,800 episodes first detected between 10th Oct 2020 and 6th Dec 2020 inclusive, with negatives bounding the start and end time of the episode.
A negative test before the start of the episode provides a lower bound on the episode's start time; otherwise, there is little information on its length.
Similar information comes from the negative test after the end of the episode.
This is typically observed, unless the individual is lost to follow-up before their episode is ended, which happens rarely.
I will incorporate episodes excluded in this step in the model for undetected infections.

@@

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$; in this case, $n_{ia} = 0$ admissible episodes occurred in individuals with test schedule $\sched_i$.
  Otherwise, if $i \in \set{D}$ then $n_{ia} =1$ admissible episode occurred in an individual with test schedule $\sched_{i}$.
  Conditional on the episode occurring in individual $i$, the episode is admissible with probability $p_{ia} = \prob((b, e) \in \alpha_i \mid \vec{\theta})$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these occur in individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, the episode is undetected with probability $p_{iu} = \prob((b, e) \in \Omega_i \mid \vec{\theta})$.
\item
  Inadmissible episodes, $\beta_i$, which are not consistent with the data observed but would have been
  detected.
  These are all remaining episodes not in the previous sets, with $n_{i\bar{a}} = 0$ of these occurred in any individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in these individuals, the episode inadmissible with probability $p_{i\bar{a}} = \prob((b, e) \in \beta_i \mid \vec{\theta})$.
\end{itemize}

@@

{\color{red} Would this be the place to introduce a new subsection? on how to calculate the likelihood? Or on line 242?}
The undetected classes for each individual cannot be distinguished by any observable data.
Each class's count needs to be marginalised over to calculate the likelihood.
Because the counts have a joint multinomial distribution, this is equivalent to only considering the sum of the counts.
The overall probability of an episode being unobserved is $\pnodet = \frac{1}{\Ncis} \sum_{i=1}^{\nsched} \eta_i p_{iu}$.
Furthermore, the inadmissible classes are guaranteed to have zero episodes.
Counts of zero do not contribute to the multinomial probability mass function (they multiply by 1).
Therefore, I combine the undetected classes into a single class and ignore the inadmissible classes.

{\color{red} subsection here instead of line 233?}
Let $\na = (n_{1a}, \dots, n_{\nsched a})^T$ be the vector of counts of episodes in the admissible classes.
The counts $\na$ and $\nnodet$ are jointly distributed multinomially.
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{\nsched} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
p(\na \mid \ntot, \theta)
&= p(\nnodet, \na\mid \ntot, \theta) \\
&= \frac{\ntot!}{\nnodet! \prod_{i=1}^{\nsched} n_{ia}!}  \pnodet^{\nnodet}\prod_{i=1}^{\nsched} \left(\frac{1}{\nsched}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $\nnodet = \ntot - \ndet$, $n_{ia} = 1$ for $i \in \set{D}$, and $n_{ia} = 0$ otherwise.

@@

To start, I derive an analytical solution to $\sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet}$ under the prior $\ntot \dist \NBc(\mu, r)$.
I assume that $\theta$, which are only the parameters of the survival distribution, has a prior independent of $\ntot$, the number of infections that occurred in the cohort.
Therefore, $p(\ntot \mid \theta) = p(\ntot)$.
Putting a negative binomial prior on $\ntot$ is equivalent to the following gamma-Poisson composite; its use simplifies the derivation.
\begin{align}
\ntot \mid \lambda &\dist \Poi(\lambda) \\
\lambda &\dist \GamDist(a, b)
\end{align}
where $b = r / \mu$ and $a = r$.
Hence:
\begin{align}
&\sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} p(\ntot) \\
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} p(\ntot \mid \lambda) p(\lambda) d\lambda &\text{$\lambda$ explicit}\\
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \frac{\lambda^{\ntot} e^{-\lambda}}{\ntot!} p(\lambda) d\lambda &\ntot \dist \Poi\\
%&= \int \sum_{\ntot=\ndet}^\infty \frac{1}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \lambda^{\ntot-\ndet} \lambda^{\ndet} e^{-\lambda} p(\lambda) d\lambda \\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) \sum_{\nnodet=0}^\infty \frac{(\pnodet \lambda)^{\nnodet}}{\nnodet!} d\lambda &\nnodet = \ntot-\ndet\\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) e^{\lambda \pnodet} d\lambda &\text{Maclaurin series of $e$} \\
&= \int \lambda^{\ndet} e^{-\lambda(1 - \pnodet)} \frac{b^a}{\Gamma(a)} \lambda^{a-1} e^{-b\lambda} \lambda d\lambda &\text{Gamma pdf}\\
&= \int \frac{b^a}{\Gamma(a)} \lambda^{a+\ndet-1} e^{-(b+1-\pnodet)\lambda} \lambda d\lambda \\
&= \frac{b^a}{\Gamma(a)} \frac{\Gamma(a+\ndet)}{(b+1-\pnodet)^{a+\ndet}} &\text{Gamma pdf}\\
&\propto (b+1-\pnodet)^{-(a+\ndet)} \\
&= (r/\mu + 1 - \pnodet)^{-(r+\ndet)} &\text{sub in $\mu$ and $r$}\\
&\propto(r + \mu (1- \pnodet))^{-(r+\ndet)}.
\end{align}

Substituting this into \cref{perf-test:eq:full-posterior} gives:
\begin{align}
p(\vec{\theta} \mid \na)
&\propto p(\vec{\theta}) \left( \prod_{i \in \set{D}} p_{ia} \right) (r + \mu (1- \pnodet))^{-(r+\ndet)} \label{perf-test:eq:full-posterior-simplified}.
\end{align}

% $\ntot$ is generally a nuisance parameter, however, its posterior distribution can be useful for diagnostic purposes (as in X\todo{ref where I use this in the next chapter}).
% Its posterior can be reconstructed using a posterior sample of $\vec{\theta}$.
% For each posterior sample of $\vec{\theta}$, sample from the full conditional of $\ntot$ to give the joint posterior of $\ntot$ and $\vec{\theta}$.
% The full conditional of $\ntot$ is given by:
% \begin{align}
% p(n_\text{tot} \mid \ncis, \vec{\theta})
% &\propto p(\ntot \mid \vec{\theta}) p(\ncis \mid \vec{\theta}, \ntot) \\
% &\propto p(\ntot \mid \vec{\theta}) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} &\text{by \cref{perf-test:eq:augmented-likelihood}} \\
% &\propto \frac{\Gamma(r + \ntot)}{\ntot!} \left( \frac{\mu}{r + \mu} \right)^{\ntot} \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot} \\
% &= \frac{\Gamma(r + \ntot)}{(\ntot-\ndet)!} \left( \frac{\mu \pnodet}{r + \mu} \right)^{\ntot}  \\
% &= \frac{\Gamma((r + \ndet) + (\ntot- \ndet))}{(\ntot-\ndet)!} \left( \frac{\mu \pnodet}{r + \mu} \right)^{\ntot-\ndet}.
% \end{align}
% Comparing this expression to the pmf of a negative binomial, we find that $\nnodet = \ntot - \ndet$ is distributed negative binomial with size parameter $r+\ndet$ and probability parameter $(r + \mu(1 - \pnodet))/(r + \mu)$.
%The mean of this distribution is $(r+\ndet)\mu \pnodet/(r+\mu(1-\pnodet))$.

Next, I derive $p_{ia}$.
By definition (see~\cpageref{perf-test:def:admissible}), we have, for any $i \in \set{D}$:
\begin{align}
p_{ia} =& \prob((b, e) \in \alpha_i \mid \vec{\theta}) \\
\alpha_i =& \{ (b, e) : l_{j(i)}^{(b)} \leq b \leq r_{j(i)}^{(b)} \wedge l_{j(i)}^{(e)} \leq e \leq r_{j(i)}^{(e)}\}.
\intertext{Omitting the conditioning on $\vec{\theta}$, this gives:}
p_{ia}
=& \prob \left( l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)}, l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \right) \\
=& \prob \left( l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \mid l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)} \right) \\
   &\times\prob \left( l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)} \right) \\
=& \sum_{b = l_{j(i)}^{(l)}}^{r_{j(i)}^{(b)}} \prob \left( l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \mid B_{j(i)} = b \right) \prob \left(B_{j(i)} = b \right) \\
=& \sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \prob \left( l_{j(i)}^{(e)} - b + 1 \leq D_{j(i)} \leq r_{j(i)}^{(e)} - b + 1 \right) \prob \left(B_{j(i)} = b \right) &\text{by def of $D_{j(i)}$} \\
=& \sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \left( S_{\vec{\theta}}(l_{j(i)}^{(e)} - b + 1) - S_{\vec{\theta}}(r_{j(i)}^{(e)} - b + 1) \right) \prob \left(B_{j(i)} = b \right) &\text{by def of $S_{\vec{\theta}}$} \\
\propto& \sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \left( S_{\vec{\theta}}(l_{j(i)}^{(e)} - b + 1) - S_{\vec{\theta}}(r_{j(i)}^{(e)} - b + 1) \right)
\end{align}
under the assumption of uniform probability of infection time.
This is the standard form of the likelihood for doubly interval censored data with truncation.

The next step is to derive $1 - p_u = 1 - \frac{1}{N} \sum_{i=1}^{\nsched} p_{iu} = \frac{1}{N} \sum_{i=1}^{\nsched} (1 - p_{iu})$.
Therefore, the crucial component is $1 - p_{iu} = 1 - \prob((B_{j(i)}, E_{j(i)}) \in \Omega_i \mid \vec{\theta})$, one minus the probability that episode $i$ was undetected.
An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t.
\end{align}
Then $\Omega_i$ can be written as:
\begin{align}
\Omega_i = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
\end{align}
Therefore, suppressing the conditioning on $\vec{\theta}$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_{j(i)}, E_{j(i)}) \in \Omega_i) \\
&= 1 - \prob( \tau_{\sched_i}(B_{j(i)}) + B_{j(i)} > E_{j(i)} \vee B_{j(i)} \leq \min(\sched_i) ) \\
&= 1 - \prob( \tau_{\sched_i}(B_{j(i)}) + 1 > D_i , B_{j(i)} > \min(\sched_i) )  - \prob( B_{j(i)} \leq \min(\sched_i) ) \\
&= 1 - \sum_{b={\min\sched_i} + 1}^T \prob( \tau_{\sched_i}(b) + 1> D_i) \prob(B_{j(i)} = b) - \sum_{t=1}^{\min\sched_i} \prob(B_{j(i)} = b)\\
&= 1 - \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T (1 - S_{\vec{\theta}}(\tau_{\sched_i}(b) + 1)) - \frac{1}{T} \min\sched_i \\
&= \frac{1}{T} \left(T - \sum_{b=\min\sched_i+1}^T (1 - S_{\vec{\theta}}(\tau_{\sched_i}(b) + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \left(T - (T - \min\sched_i) + \sum_{b=\min\sched_i+1}^T S_{\vec{\theta}}(\tau_{\sched_i}(b + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T S_{\vec{\theta}}(\tau_{\sched_i}(b) + 1). \label{perf-test:eq:piu}
\end{align}

@@

In this section I specify the form of $S_{\vec{\theta}}(t)$.
% As frequently observed in the literature~\autocite[e.g.][]{heBayesian}, it is more convenient to operate with the hazard than the survival or probability mass function directly.
% The mass function must sum to one.
% The survival must be monotonically decreasing
% But, there are no constraints across multiple hazards.
% This would complicate the inference process in a way that cannot be removed through a simple transformation.
% The only constraint on the hazard is to be in the interval $[0, 1]$, since they are probabilities.
% The [0, 1] interval is easily be mapped to the unconstrained reals using a logit transformation (where $\logit(x) = \log(x/(1-x))$).
Priors cannot be vague on both the hazard and the survival, the implications of which I consider in what follows.
Informative priors are also attractive to allow the incorporation of the estimates from \cref{E-ATACCC}.

@@

Instead, I propose a weakly informative (or vague) prior Beta(0.1, 1.9), which has mean 0.05 and minimal information.
The amount of information in a Beta distribution is related to the sum of its parameters.
Here, the sum equals 2, the same as the flat prior case.
The central 95\% probability mass of Beta(0.1, 1.9) is 0.00--0.47.
The central estimate, of 0.05, is in line with previous estimates that the median duration i in the range 15--20 days~\autocite{cevikShedding}.
This prior gives a very vague prior predictive distribution on the survival time (see \cref{perf-test:fig:vague-prior}).
\begin{figure}
  \centering \includegraphics{cis-perfect-testing/vague-prior}
  \caption[Vague prior for the hazard]{
    (A) Density of Beta(0.1, 1.9) vague prior.
    (B) Prior predictive survival time (mean and 95\% CrI).
    Note the different x-axis to \cref{perf-test:fig:flat-prior}.
  }
  \label{perf-test:fig:vague-prior}
\end{figure}

@@

Smoothing priors encode the biological consideration that the hazard should not vary significantly from day-to-day.
It is the Bayesian equivalent of a penalized likelihood, used frequently in the non-parametric frequentist setting~\autocite[e.g.][]{bacchettiNonparametric}.
Many possible forms of smoothing priors exist, including splines, Gaussian Processes, and random walks.
I use a second-order random walk (RW2) because it is simple and produces sensible prior predictive distributions.

>>>>

% \section{Definitions}
% \todo[inline]{These will be defined in the introductory chapters, but are here so that this draft is understandable}
% \begin{itemize}
% \item
%   We assume that, for all episodes $i$, $D_i$ is iid and independent
%   of the time of the infection. Define the survival function
%   $\prob(D_i \geq t \mid B_{j(i)} = b, \theta) = \prob(D_i \geq t \mid \theta) = S_\theta(t)$,
%   where $\theta$ are the parameters controlling the survival
%   distribution (the discussion in this document is valid regardless of
%   the model specified for $S_\theta$ and hence we consider $\theta$
%   as an arbitrary vector of parameters).
% \item
%   The beginning of episode $i$ is known to occur in the interval
%   $[l_{j(i)}^{(b)}, r_{j(i)}^{(b)}]$, and similarly for the end of the infection
%   in $[l_{j(i)}^{(e)}, r_{j(i)}^{(e)}]$.
% \end{itemize}

@@

Doubly interval censored and truncated data are common when assessing the survival time of bird and insect nests~\autocite{heiseyABCs}.
\textcite{heiseyModelling}, building on on long-standing methodology~\autocite{dempsterMaximum,turnbullEmpirical}, develops a likelihood-based inference framework, allowing for arbitrary truncation patterns and double interval censoring.
They apply it to a situation where the visit times (possible detection times) are common to all the individuals, simplifying the likelihood.
I adapt this framework for CIS, when each individual has their own visit schedule (see section \cref{perf-test:sec:model}).

@@

The model will be used to analyse on the 4,800 episodes first detected between 10th Oct 2020 and 6th Dec 2020 inclusive, with negatives bounding the start and end time of the episode.
A negative test before the start of the episode provides a lower bound on the episode's start time; otherwise, there is little information on its length.
Similar information comes from the negative test after the end of the episode.
This is typically observed, unless the individual is lost to follow-up before their episode is ended, which happens rarely.
I will incorporate episodes excluded in this step in the model for undetected infections. {\color{red} introduce here what is being done in teh next subsections to give some ideas of why you do what you do}

@@

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$; in this case, $n_{ia} = 0$ admissible episodes occurred in individuals with test schedule $\sched_i$.
  Otherwise, if $i \in \set{D}$ then $n_{ia} =1$ admissible episode occurred in an individual with test schedule $\sched_{i}$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, the episode is admissible with probability $p_{ia} = \prob((b, e) \in \alpha_i \mid \theta)$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these occur in individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, the episode is undetected with probability $p_{iu} = \prob((b, e) \in \Omega_i \mid \theta)$.
\item
  Inadmissible episodes, $\beta_i$, which are not consistent with the data observed but would have been
  detected.
  These are all remaining episodes not in the previous sets, with $n_{i\bar{a}} = 0$ of these occurred in any individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in these individuals, the episode inadmissible with probability $p_{i\bar{a}} = \prob((b, e) \in \beta_i \mid \theta)$.
\end{itemize}

@@

{\color{red} Would this be the place to introduce a new subsection? on how to calculate the likelihood? }
The undetected classes for each individual cannot be distinguished by any observable data.
Each class's count needs to be marginalised over to calculate the likelihood.
Because the counts have a joint multinomial distribution, this is equivalent to only considering the sum of the counts.
The overall probability of an episode being unobserved is $\pnodet = \frac{1}{\Ncis} \sum_{i=1}^{\nsched} \eta_i p_{iu}$.
Furthermore, the inadmissible classes are guaranteed to have zero episodes.
Counts of zero do not contribute to the multinomial probability mass function (they multiply by 1).
Therefore, I combine the undetected classes into a single class and ignore the inadmissible classes.

{\color{red} subsection on likelihood calculation here instead of above?}
Let $\na = (n_{1a}, \dots, n_{\nsched a})^T$ be the vector of counts of episodes in the admissible classes.
The counts $\na$ and $\nnodet$ are jointly distributed multinomially.
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{\nsched} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
p(\na \mid \ntot, \theta)
&= p(\nnodet, \na\mid \ntot, \theta) \\
&= \frac{\ntot!}{\nnodet! \prod_{i=1}^{\nsched} n_{ia}!}  \pnodet^{\nnodet}\prod_{i=1}^{\nsched} \left(\frac{1}{\nsched}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $\nnodet = \ntot - \ndet$, $n_{ia} = 1$ for $i \in \set{D}$, and $n_{ia} = 0$ otherwise.

@@

\paragraph{name of the sum} To start, I derive an analytical solution to $\sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet}$ under the prior $\ntot \dist \NBc(\mu, r)$.
I assume that $\theta$, which are only the parameters of the survival distribution, has a prior independent of $\ntot$, the number of infections that occurred in the cohort.
Therefore, $p(\ntot \mid \theta) = p(\ntot)$.
Putting a negative binomial prior on $\ntot$ is equivalent to the following gamma-Poisson composite; its use simplifies the derivation.
\begin{align}
\ntot \mid \lambda &\dist \Poi(\lambda) \\
\lambda &\dist \GamDist(a, b)
\end{align}
where $b = r / \mu$ and $a = r$.
Hence:
\begin{align}
&\sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} p(\ntot) \\
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} p(\ntot \mid \lambda) p(\lambda) d\lambda &\text{$\lambda$ explicit}\\
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \frac{\lambda^{\ntot} e^{-\lambda}}{\ntot!} p(\lambda) d\lambda &\ntot \dist \Poi\\
%&= \int \sum_{\ntot=\ndet}^\infty \frac{1}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \lambda^{\ntot-\ndet} \lambda^{\ndet} e^{-\lambda} p(\lambda) d\lambda \\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) \sum_{\nnodet=0}^\infty \frac{(\pnodet \lambda)^{\nnodet}}{\nnodet!} d\lambda &\nnodet = \ntot-\ndet\\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) e^{\lambda \pnodet} d\lambda &\text{Maclaurin series of $e$} \\
&= \int \lambda^{\ndet} e^{-\lambda(1 - \pnodet)} \frac{b^a}{\Gamma(a)} \lambda^{a-1} e^{-b\lambda} \lambda d\lambda &\text{Gamma pdf}\\
&= \int \frac{b^a}{\Gamma(a)} \lambda^{a+\ndet-1} e^{-(b+1-\pnodet)\lambda} \lambda d\lambda \\
&= \frac{b^a}{\Gamma(a)} \frac{\Gamma(a+\ndet)}{(b+1-\pnodet)^{a+\ndet}} &\text{Gamma pdf}\\
&\propto (b+1-\pnodet)^{-(a+\ndet)} \\
&= (r/\mu + 1 - \pnodet)^{-(r+\ndet)} &\text{sub in $\mu$ and $r$}\\
&\propto(r + \mu (1- \pnodet))^{-(r+\ndet)}.
\end{align}

Substituting this into \cref{perf-test:eq:full-posterior} gives:
\begin{align}
p(\theta \mid \na)
&\propto p(\theta) \left( \prod_{i \in \set{D}} p_{ia} \right) (r + \mu (1- \pnodet))^{-(r+\ndet)} \label{perf-test:eq:full-posterior-simplified}.
\end{align}

% $\ntot$ is generally a nuisance parameter, however, its posterior distribution can be useful for diagnostic purposes (as in X\todo{ref where I use this in the next chapter}).
% Its posterior can be reconstructed using a posterior sample of $\theta$.
% For each posterior sample of $\theta$, sample from the full conditional of $\ntot$ to give the joint posterior of $\ntot$ and $\theta$.
% The full conditional of $\ntot$ is given by:
% \begin{align}
% p(n_\text{tot} \mid \ncis, \theta)
% &\propto p(\ntot \mid \theta) p(\ncis \mid \theta, \ntot) \\
% &\propto p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} &\text{by \cref{perf-test:eq:augmented-likelihood}} \\
% &\propto \frac{\Gamma(r + \ntot)}{\ntot!} \left( \frac{\mu}{r + \mu} \right)^{\ntot} \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot} \\
% &= \frac{\Gamma(r + \ntot)}{(\ntot-\ndet)!} \left( \frac{\mu \pnodet}{r + \mu} \right)^{\ntot}  \\
% &= \frac{\Gamma((r + \ndet) + (\ntot- \ndet))}{(\ntot-\ndet)!} \left( \frac{\mu \pnodet}{r + \mu} \right)^{\ntot-\ndet}.
% \end{align}
% Comparing this expression to the pmf of a negative binomial, we find that $\nnodet = \ntot - \ndet$ is distributed negative binomial with size parameter $r+\ndet$ and probability parameter $(r + \mu(1 - \pnodet))/(r + \mu)$.
%The mean of this distribution is $(r+\ndet)\mu \pnodet/(r+\mu(1-\pnodet))$.

\Paragraph{$p_{ia}$}
By definition (see~\cpageref{perf-test:def:admissible}), we have, for any $i \in \set{D}$:
\begin{align}
p_{ia} =& \prob((b, e) \in \alpha_i \mid \theta) \\
\alpha_i =& \{ (b, e) : l_{j(i)}^{(b)} \leq b \leq r_{j(i)}^{(b)} \wedge l_{j(i)}^{(e)} \leq e \leq r_{j(i)}^{(e)}\}.
\intertext{Omitting the conditioning on $\theta$, this gives:}
p_{ia}
=& \prob \left( l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)}, l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \right) \\
=& \prob \left( l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \mid l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)} \right) \\
   &\times\prob \left( l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)} \right) \\
=& \sum_{b = l_{j(i)}^{(l)}}^{r_{j(i)}^{(b)}} \prob \left( l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \mid B_{j(i)} = b \right) \prob \left(B_{j(i)} = b \right) \\
=& \sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \prob \left( l_{j(i)}^{(e)} - b + 1 \leq D_{j(i)} \leq r_{j(i)}^{(e)} - b + 1 \right) \prob \left(B_{j(i)} = b \right) &\text{by def of $D_{j(i)}$} \\
=& \sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \left( S_\theta(l_{j(i)}^{(e)} - b + 1) - S_\theta(r_{j(i)}^{(e)} - b + 1) \right) \prob \left(B_{j(i)} = b \right) &\text{by def of $S_\theta$} \\
\propto& \sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \left( S_\theta(l_{j(i)}^{(e)} - b + 1) - S_\theta(r_{j(i)}^{(e)} - b + 1) \right)
\end{align}
under the assumption of uniform probability of infection time.
This is the standard form of the likelihood for doubly interval censored data with truncation.

%The next step is to derive 
\paragraph{$1 - p_u$}
Note that $1 - p_u= 1 - \frac{1}{N} \sum_{i=1}^{\nsched} p_{iu} = \frac{1}{N} \sum_{i=1}^{\nsched} (1 - p_{iu})$.
Therefore, the crucial component is $1 - p_{iu} = 1 - \prob((B_{j(i)}, E_{j(i)}) \in \Omega_i \mid \theta)$, one minus the probability that episode $i$ was undetected.
An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t.
\end{align}
Then $\Omega_i$ can be written as:
\begin{align}
\Omega_i = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
\end{align}
Therefore, suppressing the conditioning on $\theta$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_{j(i)}, E_{j(i)}) \in \Omega_i) \\
&= 1 - \prob( \tau_{\sched_i}(B_{j(i)}) + B_{j(i)} > E_{j(i)} \vee B_{j(i)} \leq \min(\sched_i) ) \\
&= 1 - \prob( \tau_{\sched_i}(B_{j(i)}) + 1 > D_i , B_{j(i)} > \min(\sched_i) )  - \prob( B_{j(i)} \leq \min(\sched_i) ) \\
&= 1 - \sum_{b={\min\sched_i} + 1}^T \prob( \tau_{\sched_i}(b) + 1> D_i) \prob(B_{j(i)} = b) - \sum_{t=1}^{\min\sched_i} \prob(B_{j(i)} = b)\\
&= 1 - \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \frac{1}{T} \min\sched_i \\
&= \frac{1}{T} \left(T - \sum_{b=\min\sched_i+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \left(T - (T - \min\sched_i) + \sum_{b=\min\sched_i+1}^T S_\theta(\tau_{\sched_i}(b + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T S_\theta(\tau_{\sched_i}(b) + 1). \label{perf-test:eq:piu}
\end{align}

@@

In this section I specify the form of $S_\theta(t)$.
I parameterise $S$ in terms of the hazard $S_{\vec{\theta}}(t) = \prod_{i=1}^{t-1} (1 - \lambda(i))$ (see \cref{E-inc-prev:sec:definitions}).
Therefore, $\vec{\theta} = [\lambda_1, \dots, \lambda_{\dmax-1}]^T$.

Priors cannot be vague on both the hazard and the survival, the implications of which I consider in what follows.
Informative priors are also attractive to allow the incorporation of the estimates from \cref{E-ATACCC}.

@@

Instead, I propose a weakly informative (or vague) prior Beta(0.1, 1.9), which has mean 0.05 and minimal information.
The amount of information in a Beta distribution is related to the sum of its parameters.
Here, the sum equals 2, the same as the flat prior case.
The central 95\% probability mass of Beta(0.1, 1.9) is 0.00--0.47.
The central estimate, of 0.05, is in line with previous estimates that the median duration in the range 15--20 days~\autocite{cevikShedding}.
This prior gives a very vague prior predictive distribution on the survival time (see \cref{perf-test:fig:vague-prior}).
\begin{figure}
  \centering \includegraphics{cis-perfect-testing/vague-prior}
  \caption[Vague prior for the hazard]{
    (A) Density of Beta(0.1, 1.9) vague prior.
    (B) Prior predictive survival time (mean and 95\% CrI).
    Note the different x-axis to \cref{perf-test:fig:flat-prior}.
  }
  \label{perf-test:fig:vague-prior}
\end{figure}

@@

Smoothing priors encode the biological consideration that the hazard should not vary significantly from day-to-day.
It is the Bayesian equivalent of a penalized likelihood, which is used frequently in the non-parametric frequentist setting~\autocite[e.g.][]{bacchettiNonparametric}.
Many possible forms of smoothing priors exist, including splines, Gaussian Processes, and random walks.
I use a second-order random walk (RW2) because it is simple and produces sensible prior predictive distributions.
